Exception Handling Bug Hazards in Android

### 2016-04-18 SPRINGER_JOURNAL/msr_journal.tex
addressing journal reviewers comments - REVIWER ***2***

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
 \emph{explicitly thrown} runtime exceptions 
should be documented as part of the \emph{exception interface} of the reusable
methods of libraries/frameworks. To investigate the conformance to this practice, we first filtered out all the exceptions implicitly
 signaled by  the runtime environment (due to programming mistakes) -- since these exceptions 
should not be documented in the method signature.  Then we inspected
the code for each method 
(defined either in the Android  Application Framework or in third-party libraries) 
explicitly signaling a runtime exception. 
Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We found 79 methods (both from libraries and the Android platform) that  explicitly threw a runtime exception 
without listing it in the \emph{exception interface} (i.e., using 
a throws clause in the method signature). From this set only one method (defined in a library)
included an @throws tag in its Javadoc -- reporting that the given runtime exception
could be thrown in some conditions. These methods were responsible for 
267 exception stack traces mined in this study.

@@

\emph{\textbf{Runtime Exception wrapping an Error.}} 
From Table~\ref{tab:wrappingandroid}, we see that most of these wrappings	
are performed by the Android platform (50.7\%).
The code snippet below was extracted from Android and shows a general catch clause 
that converts any instance of Throwable (signaled during the execution
of an asynchronous task) into an instance of RuntimeException and re-throws it.  
Table~\ref{tab:exampeswrap} presents examples of exceptions that were actually wrapped in this code snippet.  
Such wrappings mask an unrecoverable Error into a general runtime exception.

@@

\emph{\textbf{Runtime Exception wrapping a Checked Exception.}} 
This wrapping was responsible for 49.5\% of
the cross-type wrappings.  From this set 50\%  were performed by methods defined by the Android platform. 
We observe that this is a common implementation practice in the methods of the Android platform.
However, using such a general exception is considered a bad practice
according to the Java specification and common guidelines, as it loses contextual information about the exception.

@@

\emph{\textbf{Checked Exception wrapping an Error.}} Most of these wrappings
 were also caused by the reflection library used by applications' methods. The methods responsible for the wrappings
were also native methods written in C. Table~\ref{tab:exampeswrap} illustrates
some of these wrappings -- some of them are masking an OutOfMemoryError
into a checked exception. Such wrappings may also mask an unrecoverable error
and may lead to the ``exception confusion'' described next.

@@

Some surveys focused on Android developers such as xxxx.

@@

Kochhar et al conducted a survey-based study whose goal was to discover the commonly used tools for mobile app testing and the problems faced by developers while testing the apps. The survey was sent to 3,905 emails of Android developers and received 83 responses (response rate of 2.13%).

>>>>

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
 \emph{explicitly thrown} runtime exceptions 
should be documented as part of the \emph{exception interface} of the reusable
methods of libraries/frameworks. To investigate the conformance to this practice, we first filtered out all the exceptions implicitly
 signaled by  the runtime environment (due to programming mistakes) -- since these exceptions 
should not be documented in the method signature\footnote{The filtering was performed in two steps:
firstly, we analyzed the bytecode of the JVM and identified all runtime exceptions defined by it (e.g.,
java.lang.NullPointerExceptoin,java.lang.ArrayIndexOutOfBounds), then
if the libraries/framework method was signaling one of such exceptions it was filtered out from the analysis}.
 Then we inspected
the code for each method 
(defined either in the Android  Application Framework or in third-party libraries) 
explicitly signaling a runtime exception. 
Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We found 79 methods (both from libraries and the Android platform) that  explicitly threw a runtime exception 
without listing it in the \emph{exception interface} (i.e., using 
a throws clause in the method signature). From this set only one method (defined in a library)
included an @throws tag in its Javadoc -- reporting that the given runtime exception
could be thrown in some conditions. These methods were responsible for 
267 exception stack traces mined in this study.

@@

\emph{\textbf{Runtime Exception wrapping a Checked Exception.}} 
This wrapping was responsible for 49.5\% of
the cross-type wrappings.  From this set 50\%  were performed by methods defined by the Android platform. 
We observe that this is a common implementation practice in the methods of the Android platform.
According to Java best practices checked exceptions represent conditions from which the caller
is expected to recover. By converting a checked exception in a general Runtime class,
besides loosing contextual information about the exception,
the client can simply ignore such exception - which can negatively impact the app robustness. 
\textcolor{red}{Altough it is considered a bad practice, when such wrapping is accompained with a well defined 
exception handling policy (which assures that a proper handling will be provided within the app)
the impact on app robustness can be mitigated.}

\emph{\textbf{Runtime Exception wrapping an Error.}} 
From Table~\ref{tab:wrappingandroid}, we see that most of these wrappings	
are performed by the Android platform (50.7\%).
The code snippet below was extracted from Android and shows a general catch clause 
that converts any instance of Throwable (signaled during the execution
of an asynchronous task) into an instance of RuntimeException and re-throws it.  
Table~\ref{tab:exampeswrap} presents examples of exceptions that 
were actually wrapped in this code snippet: java.lang.RuntimeException wrapped
an java.lang.OutOfMemoryError and a java.lang.StackOverflowError.

\textcolor{red}{Such Errors result from failures detected by the runtime environment which indicate 
resource deficiencies from which the program cannot possibly recover.
Hence, `hiding" an unrecoverable condition into a runtime exception
can lead to a scenario where the developer tries to handle such exception, leading the program
to na unpredictable state. This kind of wrapping should be avoided, since represent a
serious bug hazard.}

@@

@@

\emph{\textbf{Checked Exception wrapping an Error.}} Most of these wrappings
 were also caused by the reflection library used by applications' methods. The methods responsible for the wrappings
were also native methods written in C. Table~\ref{tab:exampeswrap} illustrates
some of these wrappings -- some of them are masking an OutOfMemoryError
into a checked exception. 
\textcolor{red}{By confronting the API user with a checked exception,
the API designer is forcing the client to handle the exceptional condition. 
But in this case the exception condition in and unrecoverable error,
which makes infeasible the task of providing a proper handler. 
This kind of wrapping is even more dangerous the previous one
and may lead to the ``exception confusion'' described next.}

@@

Some of such exploratory surveys were applied to Android developers  [][][][]. Despite the low response rate
of all these surverys (between 4\% and 2\%), these studies could get interesting insights about the state 
of practice of Android development. Kochhar et al conducted a survey-based study to discover the 
commonly used tools for mobile app testing as well as the problems 
faced by developers while testing the apps. The survey was sent to 3,905 emails of Android developers 
and received 83 responses (response rate of 2.13%). 
Joorabchi et al. performed a survey-based study whose goal was to better understand the the main challenges 
developers face when building apps for different mobile devices. 
They interviewed 12 senior mobile developers and performed a semi-structured survey, with 188 respondents
 from the mobile development community - since the survey was shared via e-mail groups and social media websites
response rate of the survey could not be calculated. Linares-Vásquez et al. surveyed 485 open source Android 
app and library developers (hosted on GitHub)  to understand developers’ practices for detecting and fixing 
performance bottlenecks in mobile apps.  This work emailed the survey to 24,340 developers and received 
628 responses - the response rate was 2.6\%.  
Bavota at al. investigated the impact of API change-proness and fault-proneness on the user ratings of Android apps. 
They surveyed developers to assess their perspective on whether such problems could be the cause for unfavorable user ratings. 
The response rate was 4\%.
None of these surveys, however, assed the how developers deal with exceptions in Android nor 
the developers' perspective regarding a set of exception handling bug hazards. Our survey
was the first exploratory study focusing on exception handling issues in Android development.

@@

Kochhar et al conducted a survey-based study whose goal was to discover the commonly used tools for mobile app testing as well as the problems faced by developers while testing the apps. The survey was sent to 3,905 emails of Android developers and received 83 responses (response rate of 2.13%).
